Joint Likelihood Deconvolution of Astronomical Images in the Presence of Poisson Noise

### 2024-02-16 src/tex/ms.tex
[showyourwork] overleaf sync

    \subsection{Results}
    \subsubsection{Comparing Methods}
    %
    \begin{figure*}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/comparison-scenarios.pdf}
            \caption{
                Comparison of the different deconvolution methods and prior assumptions for selected source scenarios, as described in Section~\ref{subsec:test-datasets}. The simulations depicted used a fixed background level of $\lambda_{Bkg} = \qty[mode = math]{0.01}{cts/pix}$, a uniform exposure, and a Gaussian PSF of width $\sigma_{PSF} = \qty[mode = math]{2}{pix}$, corresponding to the \enquote{Chandra} instrument scenario. The leftmost column shows the simulated counts and the second column shows the underlying ground truth. The remaining columns show the reconstructed flux for each method and prior assumptions for \jolideconospace. To enhance weak structures the images use an $\arcsinh$ stretching with a scale parameter of $a=0.02$. The stretching is the same for all images. The methods are described in detail in Section~\ref{subsec:methods}. A more detailed representation of the results is available at \url{https://jolideco.github.io/jolideco-comparison}.
            }
            \label{fig:comparison-scenarios}
            \script{comparison-scenarios.py}
        \end{centering}
    \end{figure*}
    %
    First we compare the performance of the five methods and their differing prior assumptions on the different source scenarios. Figure \ref{fig:comparison-scenarios} shows the results for a selected subset of scenarios, i.e., on A1, B3, C3 and D4.

    ]
    \begin{table*}
    \centering
        \begin{tabular}{ c|c|c|c|c|c } 
            \input{output/metrics-table.tex}
        \end{tabular}
        \label{tab:metrics}
        \script{metrics-table.py}
        \caption{Image metrics for the difference source scenarios and methods shown in Figure~\ref{fig:comparison-scenarios}. The first number in a cell shows the structural similarity index (the closer to one the better ) and the second number shows the normalized root mean square (the lower the better).}
    \end{table*}

@@

    For scenario B3 all methods recover the point sources well.  There is, however, a large difference  in the reconstruction of the Gaussian shaped extended sources. As expected, the uniform prior assumption lets the emission decompose into many point sources, neglecting the spatial correlations among them. Pylira shows slightly improved results with the gaps between the bright emission peeks being more filled. However, Pylira lacks a convincing result of reconstructing the source into a single object. The \jolideco method with the patch priors shows the best results both visually and as measured by the SSI \dvd{define SSI}. With the GLEAM v0.1 patch prior the reconstruction is almost perfect, visually almost indistinguishable from the ground truth. The Zoran \& Weiss prior also shows good reconstruction of the extended sources, but less smooth and leaving a more visible extension in the surrounding point sources.

    The weak extended source with four embedded brighter point sources of Scenario C3 is challenging for all the methods. While the methods again reconstruct the point sources well, the visual differences are large for the weak underlying extended emission. The uniform prior with 1000 iterations decomposes the extended source into a grid of almost equidistant point sources resembling only the outer boundary of the original source. Pylira shows a similar tendency but improves the reconstruction of the extended source by leaving more emission in between the bright emission spots. The best results are achieved by the patch priors, which reconstruct a plausible extended source with approximately uniform brightness. The jet like features are decomposed into a \enquote{trail} of point sources by the uniform prior and Pylira method. The patch prior achieves a much more plausible reconstruction with almost uniform brightness for the linear structures. We also note that the patch prior does not show a preferred direction or bias in reconstruction, that could be possibly encoded in the training data.

>>>>

    \subsection{Results}
    \subsubsection{Comparing Methods}
    %
    \begin{figure*}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/comparison-scenarios.pdf}
            \caption{
                Comparison of the different deconvolution methods and prior assumptions for selected source scenarios, as described in Section~\ref{subsec:test-datasets}. The simulations depicted used a fixed background level of $\lambda_{Bkg} = \qty[mode = math]{0.01}{cts/pix}$, a uniform exposure, and a Gaussian PSF of width $\sigma_{PSF} = \qty[mode = math]{2}{pix}$, corresponding to the \enquote{Chandra} instrument scenario. The leftmost column shows the simulated counts and the second column shows the underlying ground truth. The remaining columns show the reconstructed flux for each method and prior assumptions for \jolideconospace. To enhance weak structures the images use an $\arcsinh$ stretching with a scale parameter of $a=0.02$. The stretching is the same for all images. The methods are described in detail in Section~\ref{subsec:methods}. A more detailed representation of the results is available at \url{https://jolideco.github.io/jolideco-comparison}.
            }
            \label{fig:comparison-scenarios}
            \script{comparison-scenarios.py}
        \end{centering}
    \end{figure*}
    %
    First we compare the performance of the five methods and their differing prior assumptions on the different source scenarios. Figure~\ref{fig:comparison-scenarios} shows the results for a selected subset of scenarios A1, B3, C3 and D4. For quantitative comparison we list the corresponding values of the structural similarity index (SSI, \cite{Wang2004}) and the normalized root mean-squared error (NRMSE) in Table~\ref{tab:metrics}.

@@

    For scenario B3 all methods recover the point sources well.  There is, however, a large difference  in the reconstruction of the Gaussian shaped extended sources. As expected, the uniform prior assumption lets the emission decompose into many point sources, neglecting the spatial correlations among them. Pylira shows slightly improved results with the gaps between the bright emission peeks being more filled. However, Pylira lacks a convincing result of reconstructing the source into a single object. The \jolideco method with the patch priors shows the best results both visually and as measured by the SSI With the GLEAM v0.1 patch prior the reconstruction is almost perfect, visually almost indistinguishable from the ground truth. The Zoran \& Weiss prior also shows good reconstruction of the extended sources, but less smooth and leaving a more visible extension in the surrounding point sources.
    %
    \begin{table*}
    \centering
        \begin{tabular}{ c|c|c|c|c|c } 
            \input{output/metrics-table.tex}
        \end{tabular}
        \label{tab:metrics}
        \script{metrics-table.py}
        \caption{Image metrics for the difference source scenarios and methods shown in Figure~\ref{fig:comparison-scenarios}. The first number in a cell shows the structural similarity index (SSI, the closer to unity, the better) and the second number shows the normalized root mean square (NRMSE, the lower the better).}
    \end{table*}
    %
    The weak extended source with four embedded brighter point sources of Scenario C3 is challenging for all the methods. While the methods again reconstruct the point sources well, the visual differences are large for the weak underlying extended emission. The uniform prior with 1000 iterations decomposes the extended source into a grid of almost equidistant point sources resembling only the outer boundary of the original source. Pylira shows a similar tendency but improves the reconstruction of the extended source by leaving more emission in between the bright emission spots. The best results are achieved by the patch priors, which reconstruct a plausible extended source with approximately uniform brightness. The jet like features are decomposed into a \enquote{trail} of point sources by the uniform prior and Pylira method. The patch prior achieves a much more plausible reconstruction with almost uniform brightness for the linear structures. We also note that the patch prior does not show a preferred direction or bias in reconstruction, that could be possibly encoded in the training data.
